---
title: "Labor Force Participation in Indiana"
author: Kent Sellers
output: html_document
---

#### Notes

-   If N/A category is ever a notable predictor, then need to figure why the variable is N/A
-   Co-linearity problem with classifications

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaps)
library(glmnet)
library(dplyr)
library(ipumsr)
library(gridExtra)
library(ggplot2)
library(MASS)
library(tree)
library(caret)
```

#### Read the data into R

```{r Project}
ddi <- read_ipums_ddi("usa_00012.xml")
df <- read_ipums_micro(ddi)
set.seed(111)
train <- sample(1: nrow(df), 0.75*nrow(df))
```

#### Transform the data

```{r Project1, echo=FALSE}

## Recode the variables and set as factors
df$COUNTYFIP <- as.factor(df$COUNTYFIP)
df$EGR <- relevel(as.factor(recode(df$COUNTYFIP,
             '0' = '0',
             '73' = '1', '89' = '1', '91' = '1', '111' = '1', '127' = '1', '131' = '1', '149' = '1',
             '39' = '2', '49' = '2', '85' = '2', '99' = '2', '141' = '2',
             '1' = '3', '3' = '3', '33' = '3', '53' = '3', '69' = '3', '87' = '3', '113' = '3', '151' = '3', '169' = '3', '179' = '3', '183' = '3',
             '7' = '4', '15' = '4', '17' = '4', '23' = '4', '45' = '4', '67' = '4', '103' = '4', '107' = '4', '157' = '4', '159' = '4', '171' = '4', '181' = '4', 
             '11' = '5', '57' = '5', '59' = '5', '63' = '5', '81' = '5', '95' = '5', '97' = '5', '109' = '5', '145' = '5', 
             '9' = '6', '35' = '6', '41' = '6', '65' = '6', '75' = '6', '135' = '6', '139' = '6', '161' = '6', '177' = '6', 
             '21' = '7', '121' = '7', '133' = '7', '153' = '7', '165' = '7', '167' = '7', 
             '13' = '8', '27' = '8', '55' = '8', '93' = '8', '101' = '8', '105' = '8', '117' = '8', '119' = '8', 
             '5' = '9', '29' = '9', '31' = '9', '47' = '9', '71' = '9', '77' = '9', '79' = '9', '115' = '9', '137' = '9', '155' = '9',
             '19' = '10', '25' = '10', '43' = '10', '61' = '10', '143' = '10', '175' = '10', 
             '37' = '11', '51' = '11', '83' = '11', '123' = '11', '125' = '11', '129' = '11', '147' = '11', '163' = '11', '173' = '11' 
             )), ref = '1')

df$AREA <- relevel(as.factor(ifelse(df$COUNTYFIP == '0', 0,
                             ifelse(df$COUNTYFIP %in% c('25', '27', '37', '51', '55', '71', '77', '79', '83', '93', '101', '117', '123', '125', '137', '147', '155', '9', '23', '31', '41', '45', '47', '53', '65', '75', '107', '121', '135', '139', '159', '171', '177', '17', '33', '49', '69', '85', '87', '99', '103', '113', '131', '149', '151', '169', '181'), 'rural', 'metro'))), ref = 'metro')

df$RELATE <- as.factor(ifelse(df$RELATE %in% c('1'), '1', '2'))

df$AGE <- relevel(as.factor(ifelse(df$AGE %in% c('16', '17', '18', '19', '20', '21', '22', '23', '24'), '16-24',
                            ifelse(df$AGE %in% c('25', '26', '27', '28', '29', '30', '31', '32', '33', '34'), '25-34', 
                            ifelse(df$AGE %in% c('35', '36', '37', '38', '39', '40', '41', '42', '43', '44'), '35-44', 
                            ifelse(df$AGE %in% c('45', '46', '47', '48', '49', '50', '51', '52', '53', '54'), '45-54',
                            ifelse(df$AGE %in% c('55', '56', '57', '58', '59', '60', '61', '62', '63', '64'), '55-64',
                             '65+')))))), ref = '35-44')


df$HISPAN <- as.character(df$HISPAN)
df$HISPAN <- relevel(as.factor(ifelse(df$HISPAN == '0', '0',
                               '1')), ref = 1)

df$CITIZEN <- as.character(df$CITIZEN)
df$CITIZEN <- relevel(as.factor(ifelse(df$CITIZEN == '0', 'N/A',
                      ifelse(df$CITIZEN == '1', 'Y', 
                      ifelse(df$CITIZEN == '2', 'Y', 
                             'N')))), ref = 'Y')

df$SPEAKENG <- as.character(df$SPEAKENG)
df$SPEAKENG <- relevel(as.factor(ifelse(df$SPEAKENG == '1', 'N',
                                  'Y')), ref = 'N')

df$EDUC <- relevel(as.factor(ifelse(df$EDUC %in% c('0', '1', '2', '3', '4', '5', '6'), 'HS or less',
                             '1+ year of college')), ref = 1)

df$DEGFIELD <- as.character(df$DEGFIELD)
df$DEGFIELD <- relevel(as.factor(ifelse(df$DEGFIELD == '00', 'N/A',
                                 ifelse(df$DEGFIELD %in% c(11,13:25,51), 'STEM',
                                 ifelse(df$DEGFIELD %in% c(15:26,48:56), 'Humanities and Social Sciences',
                                 ifelse(df$DEGFIELD %in% c(14,22,35,41,56:59,60), 'Fine Arts and Creative Industries',
                                 ifelse(df$DEGFIELD %in% c(23,32,41,61:62), 'Professional and Applied Fields', 'OTHER')))))), ref = 'Fine Arts and Creative Industries')

df$WORKEDYR <- as.character(df$WORKEDYR)
df$WORKEDYR <- relevel(as.factor(ifelse(df$WORKEDYR == '1', 'N',
                                 ifelse(df$WORKEDYR == '2', 'N', 
                                'Y'))), ref = 'Y')

df$VETDISAB <- as.character(df$VETDISAB)
df$VETDISAB <- relevel(as.factor(ifelse(df$VETDISAB == '0', '0',
                                 ifelse(df$VETDISAB == '1', '1', 
                                  '2'))), ref = '1')

df <- df %>%
mutate(LABFORCE = as.factor(LABFORCE),
       SEX = as.factor(SEX),
       MARST = as.factor(MARST),
       MARRNO = relevel(as.factor(MARRNO), ref = '1'),
       RACE = as.factor(RACE),
       HCOVANY = as.factor(HCOVANY),
       HINSCAID = as.factor(HINSCAID),
       HINSCARE = as.factor(HINSCARE),
       SCHOOL = as.factor(SCHOOL),
       DIFFREM = as.factor(DIFFREM),
       DIFFPHYS = as.factor(DIFFPHYS),
       DIFFMOB = as.factor(DIFFMOB),
       DIFFCARE = as.factor(DIFFCARE),
       DIFFSENS = as.factor(DIFFSENS),
       DIFFEYE = as.factor(DIFFEYE),
       DIFFHEAR = as.factor(DIFFHEAR),
       VETSTAT = relevel(as.factor(VETSTAT), ref = '1')
       )

## Subset the unnecessary variables
df <- subset(df, select = -c(SAMPLE, SERIAL, CBSERIAL, HHWT, CLUSTER, COUNTYFIP, STATEFIP, STRATA, GQ, PERNUM, PERWT, RELATED, RACED, HISPAND, EDUCD, VETSTATD, DEGFIELDD, OFFPOV))

## Use below clip if I want the category spelled out instead
# df1 <- df %>%
#   mutate(LABFORCE = as_factor(lbl_clean(LABFORCE)))

```

#### Exploratory data analysis

```{r Project1.1}
# Bar
ggplot(aggregate(LABFORCE ~ AREA, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = AREA, y = LABFORCE, fill = AREA)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Metro Area Status (Area)",
        x = "Area",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ EGR, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = EGR, y = LABFORCE, fill = EGR)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Economic Growth Region (EGR)",
        x = "EGR",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ RELATE, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = RELATE, y = LABFORCE, fill = RELATE)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Head of Household Status (RELATE)",
        x = "Relate",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ SEX, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = SEX, y = LABFORCE, fill = SEX)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Sex",
        x = "Sex",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ AGE, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = AGE, y = LABFORCE, fill = AGE)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Age",
        x = "Age",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ MARST, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = MARST, y = LABFORCE, fill = MARST)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Marital Status",
        x = "Marital Status",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ MARRNO, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = MARRNO, y = LABFORCE, fill = MARRNO)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Times Married",
        x = "Times Married",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ RACE, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = RACE, y = LABFORCE, fill = RACE)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Race",
        x = "Race",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ HISPAN, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = HISPAN, y = LABFORCE, fill = HISPAN)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Hispanic Origin",
        x = "Hispanic Origin",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ CITIZEN, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = CITIZEN, y = LABFORCE, fill = CITIZEN)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Citizenship Status",
        x = "Citizenship Status",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ SPEAKENG, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = SPEAKENG, y = LABFORCE, fill = SPEAKENG)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Speaks English",
        x = "Speaks English",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ HCOVANY, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = HCOVANY, y = LABFORCE, fill = HCOVANY)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Any Health Insurance Coverage",
        x = "Any Health Insurance Coverage",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ HINSCAID, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = HINSCAID, y = LABFORCE, fill = HINSCAID)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Medicaid Status",
        x = "Medicaid Status",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ HINSCARE, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = HINSCARE, y = LABFORCE, fill = HINSCARE)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Medicare Status",
        x = "Medicare Status",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ SCHOOL, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = SCHOOL, y = LABFORCE, fill = SCHOOL)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by School Attendance",
        x = "School Attendance",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ EDUC, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = EDUC, y = LABFORCE, fill = EDUC)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Educational Attainment",
        x = "Educational Attainment",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ DEGFIELD, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = DEGFIELD, y = LABFORCE, fill = DEGFIELD)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Field of Degree",
        x = "Field of Degree",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ WORKEDYR, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = WORKEDYR, y = LABFORCE, fill = WORKEDYR)) + 
   geom_bar(stat = "identity", show.legend = FALSE) +
   labs(title = "Labor Force Participation Rate by Worked Last Year",
        x = "Worked Last Year",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ VETSTAT, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = VETSTAT, y = LABFORCE, fill = VETSTAT)) + 
   geom_bar(stat = "identity",show.legend=FALSE) +
   labs(title = "Labor Force Participation Rate by Veteran Status (VETSTAT)",
        x = "VETSTAT",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ DIFFHEAR, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = DIFFHEAR, y = LABFORCE, fill = DIFFHEAR)) + 
   geom_bar(stat = "identity",show.legend=FALSE) +
   labs(title = "Labor Force Participation Rate with Hearing Dffficulty (HEARDIFF)",
        x = "DIFFHEAR",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ DIFFEYE, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = DIFFEYE, y = LABFORCE, fill = DIFFEYE)) + 
   geom_bar(stat = "identity",show.legend=FALSE) +
   labs(title = "Labor Force Participation Rate with Vision Difficulty (DIFFEYE)",
        x = "DIFFEYE",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ DIFFSENS, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = DIFFSENS, y = LABFORCE, fill = DIFFSENS)) + 
   geom_bar(stat = "identity",show.legend=FALSE) +
   labs(title = "Labor Force Participation Rate with Hearing or Vision DIfficulty (DIFFSENS)",
        x = "DIFFSENS",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ DIFFCARE, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = DIFFCARE, y = LABFORCE, fill = DIFFCARE)) + 
   geom_bar(stat = "identity",show.legend=FALSE) +
   labs(title = "Labor Force Participation Rate with Self-Care Difficulty (DIFFCARE)",
        x = "DIFFCARE",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ DIFFMOB, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = DIFFMOB, y = LABFORCE, fill = DIFFMOB)) + 
   geom_bar(stat = "identity",show.legend=FALSE) +
   labs(title = "Labor Force Participation Rate with Independent Living Difficulty (DIFFMOB)",
        x = "DIFFMOB",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ DIFFPHYS, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = DIFFPHYS, y = LABFORCE, fill = DIFFPHYS)) + 
   geom_bar(stat = "identity",show.legend=FALSE) +
   labs(title = "Labor Force Participation Rate with Ambulatory Difficulty (DIFFPHYS)",
        x = "DIFFPHYS",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ DIFFREM, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = DIFFREM, y = LABFORCE, fill = DIFFREM)) + 
   geom_bar(stat = "identity",show.legend=FALSE) +
   labs(title = "Labor Force Participation Rate with Cognitive DIfficulty (DIFFREM)",
        x = "DIFFREM",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()
ggplot(aggregate(LABFORCE ~ VETDISAB, data = df, 
   FUN = function(x) {sum(x == 2)/length(x) * 100}), 
   aes(x = VETDISAB, y = LABFORCE, fill = VETDISAB)) + 
   geom_bar(stat = "identity",show.legend=FALSE) +
   labs(title = "Labor Force Participation Rate by Veteran Disability (VETDISAB)",
        x = "VETDISAB",
        y = "Labor Force Participation Rate (%)") +
   theme_minimal()

# Box
ggplot(df, aes(x = LABFORCE, y = YEAR, fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(INCTOT), fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(FTOTINC), fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(INCWAGE), fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(INCSS), fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(INCWELFR), fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(INCINVST), fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(INCRETIR), fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(INCSUPP), fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(INCOTHER), fill = LABFORCE)) + 
  geom_boxplot()
ggplot(df, aes(x = LABFORCE, y = log(INCEARN), fill = LABFORCE)) + 
  geom_boxplot()

# If need to put graphs in a grid to consolidate for a paper or slides
# grid.arrange(bparea, bpegr, bprelate, ncol = 2)

```

Reduce subset based on EDA

```{r Project1.3}
df <- subset(df, select = -c(AREA, CITIZEN, YEAR, INCSS, FTOTINC, INCRETIR))
```

#### Lasso

Use lasso to maximize the reduction of our subsets

```{r Project2}
x.tr <- model.matrix(LABFORCE~.,data=df)[train,-1]
x.tst <- model.matrix(LABFORCE~.,data=df)[-train,-1]
y.tr <- df[train,]$LABFORCE
y.tst <- df[-train,]$LABFORCE

lasso.cv <- cv.glmnet(x.tr, y.tr, alpha=1, family=binomial, nfold = 5)
plot(lasso.cv)
lasso <- glmnet(x.tr, y.tr, alpha=1, family=binomial, lambda=lasso.cv$lambda.1se)
print(lasso)
cbind(coef(lasso.cv, s='lambda.1se'))

lasso.pred <- predict(lasso, newx = x.tst, s = lasso.cv$lambda.1se)
lasso.class.pred <- ifelse(lasso.pred > 0.5, 1, 0)
lasso.conf <- table(y.tst, lasso.class.pred, dnn=c("TRUE", "lasso"))
print(lasso.conf)
lasso.misclass <- sum(lasso.conf[1,2],lasso.conf[2,1])/sum(lasso.conf)
print(lasso.misclass)
```

Reduce subset based on lasso

```{r Project3, echo = TRUE}
df1 <- subset(df, select = -c(HISPAN,INCTOT,DIFFSENS,DIFFEYE,DIFFHEAR))
x1 <- model.matrix(LABFORCE~.,data=df1)[,-1]
y1 <- df1$LABFORCE
```

#### Ridge

```{r Project2.2}
ridge.cv <- cv.glmnet(x.tr, y.tr, alpha=0, family=binomial, nfold = 5)
plot(ridge.cv)
ridge <- glmnet(x.tr, y.tr, alpha=0, family=binomial, lambda=ridge.cv$lambda.1se)
print(ridge)
cbind(coef(ridge.cv, s='lambda.1se'))

ridge.pred <- predict(ridge, newx = x.tst, s = ridge.cv$lambda.1se)
ridge.class.pred <- ifelse(ridge.pred > 0.5, 1, 0)
ridge.conf <- table(y.tst, ridge.class.pred, dnn=c("TRUE", "ridge"))
print(ridge.conf)
ridge.misclass <- sum(ridge.conf[1,2],ridge.conf[2,1])/sum(ridge.conf)
print(ridge.misclass)
```

### Elastic Net

```{r Project2.3}
elastic.cv <- cv.glmnet(x.tr, y.tr, alpha=0.5, family=binomial, nfold = 5)
plot(elastic.cv)
elastic <- glmnet(x.tr, y.tr, alpha=0.5, family=binomial, lambda=elastic.cv$lambda.1se)
print(elastic)
cbind(coef(elastic.cv, s='lambda.1se'))

elastic.pred <- predict(elastic, newx = x.tst, s = elastic.cv$lambda.1se)
elastic.class.pred <- ifelse(elastic.pred > 0.5, 1, 0)
elastic.conf <- table(y.tst, elastic.class.pred, dnn=c("TRUE", "elastic"))
print(elastic.conf)
elastic.misclass <- sum(elastic.conf[1,2],elastic.conf[2,1])/sum(elastic.conf)
print(elastic.misclass)
```

### LDA

Getting warning that variables are collinear.

```{r Project2.4}
lda <- lda(x, y); lda
table(y, predict(lda)$class, dnn=c("True", "LDA"))
mean(y != predict(lda)$class)
```

### QDA

```{r Project2.5}
qda <- qda(x, y); qda
table(y, predict(qda)$class, dnn=c("True", "QDA"))
mean(y != predict(qda)$class)
```

### SVM

```{r Project2.6}
svm_model <- svm(LABFORCE~., data=df, type='C')
svm_model
svm_pred<-predict(svm_model)
table(df$LABFORCE, svm_pred, dnn=c("TRUE", "SVM"))

svm_model1 <- svm(LABFORCE~., data=df1, type='C')
svm_model1
svm_pred1<-predict(svm_model1)
table(df1$LABFORCE, svm_pred, dnn=c("TRUE", "SVM"))
```

### Logistic Regression

Warning message saying "fitted probabilities numerically 0 or 1 occurred."

```{r Project2.7}
train_df <- df[train, ]
test_df <- df[-train, ]
logistic <- glm(LABFORCE ~ ., data = train_df, family = binomial)
logistic_pred <- predict(logistic, newdata = test_df, type = "response")
logistic_class_pred <- ifelse(logistic_pred > 0.5, 1, 0)
logistic.conf <- table(test_df$LABFORCE, logistic_class_pred)
print(logistic.conf)
logistic.misclass <- sum(logistic.conf[1,2],logistic.conf[2,1])/sum(logistic.conf)
print(logistic.misclass)
```

###Classification Trees

```{r Project2.11, echo = TRUE}
df2 <- as.data.frame(df)
tree.df <- tree(LABFORCE~., data=df2)
tree.df
plot(tree.df)
text(tree.df, col='red', pretty=TRUE)

tree.cv <- cv.tree(tree.df)
plot(tree.cv)

tree.prune <- prune.tree(tree.df, best=2)
tree.prune

plot(tree.prune)
text(tree.prune, pretty=TRUE, col='red')

```

# blank

```{r Project3.1, echo = TRUE}

```

